---
  format:
    revealjs: 
      theme: [default, style.scss]
      slide-number: true
      highlight-style: github
      width: 1400
      css:
       - https://cdn.jsdelivr.net/npm/bootstrap-icons@1.8.2/font/bootstrap-icons.css
---
  
##  {.center}

<h1 class="text-indigo-pink">

Analítica de Datos

<h1>

<h2>The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women</h2>

::: {style="margin-top:50px"}
### Carlos Cardona Andrade {.text-orange-gold}
:::




```{r}
#| echo: false
#| eval: true
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 4

library(tidyverse)
library(fivethirtyeight)
library(broom)


movies_data <- bechdel


movies_data <- movies_data %>%
  mutate(year_range = case_when(
    year >= 1970 & year <= 1974 ~ "1970-'74",
    year >= 1980 & year <= 1984 ~ "1980-'84",
    year >= 1990 & year <= 1994 ~ "1990-'94",
    year >= 2000 & year <= 2004 ~ "2000-'04",
    year >= 2010 & year <= 2013 ~ "2010-'13",
    TRUE ~ NA_character_  # To handle any years outside the specified ranges
  ))


movies_data <- movies_data %>%
  mutate(pass = ifelse(binary == "PASS", 1, 0),
         budget = budget/1000000,
         domgross = domgross/1000000,
         intgross = intgross/1000000)
```


## The  Alison Bechdel's test {.text-orange-gold}

A movies is going to pass the test if:



::: incremental
  1. there are at least two named women in the picture
  
  2. they have a conversation with each other at some point
  
  3. that conversation isn’t about a male character 
:::

## The data {.text-orange-gold}

1,615 films released from 1990 to 2013 with the following variables:

- whether the film passes the test or not

- which criteria the films fail to hold

- budget

- year

- domestic gross

- international gross


##  {.text-orange-gold}


```{r}
#| echo: false
#| eval: true
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 4

# Step 2: Summarize the data by year range and cleant_test
summary_data <- movies_data %>%
  group_by(year_range, clean_test) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(year_range) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(!is.na(year_range))


# Assuming `summary_data` is the data frame with year ranges, cleant_test categories, and proportions
ggplot(summary_data, aes(x = year_range, y = proportion, fill = clean_test)) +
  geom_bar(stat = "identity", color = "black", width = 0.8) +  # Stacked bars with thin black borders
  scale_fill_manual(
    values = c(
      "nowomen" = "#ff3333",           # Fewer than two women
      "notalk" = "#f26b6b",            # Women don’t talk to each other
      "men" = "#ff9999",               # Women only talk about men
      "dubious" = "lightblue",           # Dubious
      "ok" = "#169ff5"                 # Passes Bechdel Test
    ), 
    labels = c(
      "nowomen" = "Fewer than two women",
      "notalk" = "Women don’t talk to each other",
      "men" = "Women only talk about men",
      "dubious" = "Dubious",
      "ok" = "Passes Bechdel Test"
  ),
  name = ""  # Remove legend title
  ) +
  labs(
    title = "The Bechdel Test Over Time",
    subtitle = "How women are represented in movies",
    x = NULL,  # No x-axis label
    y = "Percentage"
  ) +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0, 0)) +
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.title.y = element_text(size = 12),
    panel.grid.minor = element_blank()
  ) +
  annotate("text", x = 3, y = 0.75, label = "FAIL", size = 12, color = "black", fontface = "bold") +
  annotate("text", x = 3, y = 0.25, label = "PASS", size = 12, color = "black", fontface = "bold")

```


## Does budget plays a role?  {.text-orange-gold}


```{r}
#| echo: false
#| eval: true
#| fig-align: "center"

# Summarize data to get the mean and 95% CI for each category in 'clean_test'
summary_data <- movies_data %>%
  group_by(binary) %>%
  summarise(
    mean_budget = mean(budget, na.rm = TRUE),
    se_budget = sd(budget, na.rm = TRUE) / sqrt(n()),
    ci_lower = mean_budget - 1.96 * se_budget,
    ci_upper = mean_budget + 1.96 * se_budget
  )


# Create the plot
ggplot(summary_data, aes(x = binary, y = mean_budget, fill = binary)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +  # Bar graph
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +  # Add confidence intervals
  scale_fill_manual(
    values = c(
      "PASS" = "#0066cc",  # Color for PASS
      "FAIL" = "#ff3333"   # Color for FAIL
    ),
    labels = c("PASS" = "Passes Bechdel Test", "FAIL" = "Fails Bechdel Test"),
    name = ""  # Remove legend title
  ) +
  labs(
    title = "Mean Budget by Bechdel Test Result",
    subtitle = "Including 95% Confidence Intervals",
    x = "Bechdel Test Result",
    y = "Mean Budget"
  ) +
  scale_y_continuous(labels = scales::label_dollar(), limits = c(0, 60)) +  # Set y-axis limits
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.title.y = element_text(size = 12),
    panel.grid.minor = element_blank()
  )
```




## Does any category matters? {.text-orange-gold}


```{r}
#| echo: false
#| eval: true
#| fig-align: "center"

# Summarize data to get the mean and 95% CI for each category in 'clean_test'
summary_data <- movies_data %>%
  group_by(clean_test) %>%
  summarise(
    mean_budget = mean(budget, na.rm = TRUE),
    se_budget = sd(budget, na.rm = TRUE) / sqrt(n()),
    ci_lower = mean_budget - 1.96 * se_budget,
    ci_upper = mean_budget + 1.96 * se_budget
  )


ggplot(summary_data, aes(x = clean_test, y = mean_budget, fill = clean_test)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +  # Bar graph
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +  # Add confidence intervals
  scale_fill_manual(
    values = c(
      "nowomen" = "#ff3333",
      "notalk" = "#f26b6b",
      "men" = "#ff9999",
      "dubious" = "lightblue",
      "ok" = "#0066cc"
    ),
    labels = c(
      "nowomen" = "Fewer than two women",
      "notalk" = "Women don’t talk to each other",
      "men" = "Women only talk about men",
      "dubious" = "Dubious",
      "ok" = "Passes Bechdel Test"
    ),
    name = ""  # Remove legend title
  ) +
  labs(
    title = "Mean Budget by Bechdel Test Category",
    subtitle = "Including 95% Confidence Intervals",
    x = "Bechdel Test Category",
    y = "Mean Budget"
  ) +
  scale_y_continuous(labels = scales::label_dollar()) +  # Format y-axis as currency
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray50"),
    axis.title.y = element_text(size = 12),
    panel.grid.minor = element_blank()
  )

```

## What is a regression going to tell us? {.text-orange-gold}


```{r}
#| echo: true
#| eval: true
#| fig-align: "center"
model_1 <- lm(pass ~ budget + factor(year_range), movies_data)
tidy(model_1)
```


:::{.fragment}
For each additional 1 million dollars in the budget, the probability of passing the Bechdel test decreases , on average, by 0.00186 pp keeping the other variables constant.
:::


## What is a regression going to tell us? {.text-orange-gold}


```{r}
#| echo: true
#| eval: true
#| fig-align: "center"
model_1 <- lm(pass ~ budget + factor(year_range), movies_data)
tidy(model_1)
```


:::{.fragment}
Movies from the 2000-2004 period have, on average, a 0.339 higher probability of passing the Bechdel test compared to movies from the baseline period (70-74), holding the budget constant.
:::


## What is the predictive power of our model? {.text-orange-gold}


```{r}
#| echo: true
#| eval: true
#| fig-align: "center"
glance(model_1)
```

## Are the domestic returns related to passing or failing? {.text-orange-gold}

```{r}
#| echo: true
#| eval: true
#| fig-align: "center"
model_2 <- lm(domgross ~ pass + budget, movies_data)
tidy(model_2)
```


## Are the international returns related to passing or failing? {.text-orange-gold}

```{r}
#| echo: true
#| eval: true
#| fig-align: "center"
model_3 <- lm(intgross ~ pass + budget, movies_data)
tidy(model_3)
```

## Conclusion {.text-orange-gold}

- A higher movie budget is associated with a lower likelihood of passing the Bechdel Test. This could reflect traditional gender roles in higher-budget films. (You could provide the figure if it is meaningful)

- Movies appear to have become less gender-biased over time (At least in this dataset!)

- Passing the Bechdel Test is significantly linked to domestic box office returns, but there’s no significant relationship with international returns.
